---
title: "Map"
author: "Yann Herrera Fuchs"
date: "06/10/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: "show"
---
# Install packages

```{r}
# require() is like library() except returns FALSE if missing (vs error)
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
# librarian::shelf() is like library() except installs package if missing, 
#   even from Github if include owner/repo
shelf(
  NOAA-EDAB/ecodata,
  sf)
# librarian before version 1.8.1, got error:
#   download from 'https://api.github.com/repos/NOAA-EDAB/ecodata/tarball/HEAD' failed

```

# Get spatial data

```{r}
ecodata::epu_sf

# To add a geometry list column
epu_sf <- ecodata::epu_sf %>% 
  st_transform(4326)

class(epu_sf)

# see in Environment pane, expand g1
g1 <- epu_sf$geometry[1]

plot(epu_sf)

plot(epu_sf["EPU"])

# To contextualize where this is geographically

shelf(mapview) # shelf is similar to library

mapview(epu_sf)

shelf(leaflet)

leaflet() %>% 
  #addTiles() %>% # default basemap
  addProviderTiles(providers$Esri.OceanBasemap) %>% #esri basemap
  addPolygons(data = epu_sf)
```

# Connecting to ERDAPP (NOAA server) for data extraction

```{r}
shelf(
  here, # file navigation
  rerddap) # ERDAPP r package

sst_gd_rds <- here("data/sst_gd.rds") # `here` reduces cache processing power

epu_bb <- st_bbox(epu_sf) # returns bounding coordinates of a spatial object
epu_bb

# Get info on dataset ID
sst_info <- info('jplMURSST41mday') 
sst_info

# Get data (as ERDAPP data object)
if (!file.exists(sst_gd_rds)){
  sst_gd <- griddap(
    sst_info,
    fields    = "sst",
    time      = c("2020-06-16", "2021-06-16"),
    longitude = epu_bb[c("xmin", "xmax")],
    latitude  = epu_bb[c("ymin", "ymax")])
  saveRDS(sst_gd, file = sst_gd_rds)  
}
sst_gd <- readRDS(sst_gd_rds) # Read/save file to connection and re-store under different name

sst_gd

names(sst_gd)
```

Connection is established, now we want to plot data that we extract

# Mapping and extracting data

```{r}
shelf(
  dplyr,
  ggplot2,
  mapdata)

# coastline
coast <- map_data(
  "worldHires", 
  xlim = epu_bb[c("xmin", "xmax")],
  ylim = epu_bb[c("ymin", "ymax")],
  lforce = "e")

# Select data from the last month
sst_df_last <- sst_gd$data %>%
    filter(time == max(time))

# summary(sst_last)

ggplot(
  data = sst_df_last, 
  aes(x = lon, y = lat, fill = sst)) +
  # Create a coastline polygon
  geom_polygon(
    data = coast,
    aes(x = long, y = lat, group = group), fill = "grey80") +
  # Add temperature values to each grid
  geom_tile() +
  scale_fill_gradientn(
    colors = rerddap::colors$temperature, na.value = NA) +
  theme_bw() + 
  ylab("Latitude") + 
  xlab("Longitude") +
  ggtitle("Latest SST")
```

This map shows the average SST data from the server's last month (currently June 2021), but we can extract and manipulate a time series for multiple months.

## Generating a raster stack with all-month data

```{r}
shelf(
  purrr, # for manipulating (mapping) on nested objects
  raster,
  sp,
  tidyr)
select <- dplyr::select

# Create a SST dataset
sst_tbl <- tibble(sst_gd$data) %>%
  mutate(
    # round coordinates b/c of uneven intervals, to match raster values and coordinates
    #   unique(sst_gd$data$lon) %>% sort() %>% diff() %>% table()
    #     0.0099945068359375 0.0100021362304688
    lon  = round(lon, 2),
    lat  = round(lat, 2),
    date = as.Date(time, "%Y-%m-%dT00:00:00Z")) %>%
  select(-time) %>% # remember this select is dplyr::select
  filter(!is.na(sst)) # 13M to 8.8M rows

# Create a nested table to sort coordinate data by months
sst_tbl_mo <- sst_tbl %>% 
  nest(data = c(lat, lon, sst)) %>% 
  mutate(
    raster = purrr::map(data, function(x) {
      #browser() # activating this shows us the stepwise process of the function 
      sp::coordinates(x) <- ~ lon + lat
      sp::gridded(x)     <- T # set data in grid format
      raster::raster(x) # turn x into raster format
    }))

# Convert data object to raster stack
sst_stk <- raster::stack(sst_tbl_mo$raster) 
# Define names from the monthly datasets
names(sst_stk) <- strftime(sst_tbl_mo$date, "sst_%Y.%m")
# Define coordinate reference system
raster::crs(sst_stk) <- "+proj=longlat +datum=WGS84 +no_defs"
```

## Manipulating dataset
```{r}
shelf(stringr) # for manipulating strings

# Extract monthly average from each grid in raster stack matching "epu_sf" coordinates
epu_sst_avg <- raster::extract(sst_stk, epu_sf, fun = mean, na.rm = T)
epu_sst_sd  <- raster::extract(sst_stk, epu_sf, fun = sd, na.rm = T)

epu_sst_tbl <- rbind(
  epu_sst_avg %>% 
    as_tibble() %>% 
    cbind(
      EPU  = epu_sf$EPU,
      stat = "mean") %>% 
    pivot_longer(-c(EPU, stat)),
  epu_sst_sd %>% 
    as_tibble() %>% 
    cbind(
      EPU = epu_sf$EPU,
      stat = "sd") %>% 
    pivot_longer(-c(EPU, stat))) %>% 
  mutate(
    EPU  = as.character(EPU),
    date = as.double(str_replace(name, "sst_", ""))) %>% 
  select(-name) %>% 
  pivot_wider(
    names_from  = EPU, 
    values_from = value)

# Plot timeseries
shelf(dygraphs)

epu_sst_tbl %>% 
  filter(stat == "mean") %>% 
  select(-stat) %>% 
  dygraph()
```

