---
title: "Visualize"
author: "Yann Herrera Fuchs"
date: "06/10/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: "show"
  
---

# Read data

```{r}
library(here) # for easy file navigation
library(readr)
library(DT)

# variables
url_ac <- "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_AC.csv"
# if ERDDAP server down (Error in download.file) with URL above, use this:
#    url_ac <- "https://raw.githubusercontent.com/noaa-iea/r3-train/master/data/cciea_AC.csv"
csv_ac <- here("data/cciea_AC.csv")

# download data
if (!file.exists(csv_ac))
  download.file(url_ac, csv_ac)

# read data
d_ac        <- read_csv(csv_ac, col_names = F, skip = 2)
names(d_ac) <- names(read_csv(csv_ac))

# show data
datatable(d_ac)
```

# For plotting a simple line plot of continuous data

```{r}
# Note: The data visualization ggplot cheatsheet is very helpful for this

library(dplyr)
library(ggplot2)

# subset data
d_coast <- d_ac %>% 
  # select columns
  select(time, total_fisheries_revenue_coastwide) %>% 
  # filter rows
  filter(!is.na(total_fisheries_revenue_coastwide))
datatable(d_coast)

# ggplot object
p_coast <- d_coast %>% 
  # setup aesthetics
  ggplot(aes(x = time, y = total_fisheries_revenue_coastwide)) +
  # add geometry
  geom_line()

# show plot
p_coast
```

# For plotting a trend line with two continuous variables

```{r}
p_coast + 
  geom_smooth(method = "lm")
```

# For plotting the distributions of a single continuous variable

```{r}
d_coast %>% 
  # setup aesthetics
  ggplot(aes(x = total_fisheries_revenue_coastwide)) +
  # add geometry
  geom_histogram()
```

# For plotting series with grouping variables

## Line plots
```{r}
library(stringr) # For enhanced string manipulation
library(tidyr)

d_rgn <- d_ac %>% 
  # select columns
  select(
    time, 
    starts_with("total_fisheries_revenue")) %>% 
  # exclude column
  select(-total_fisheries_revenue_coastwide) %>% 
  # pivot to longer format
  pivot_longer(-time) %>% 
  # mutate region by stripping other
  mutate(
    region = name %>% 
      str_replace("total_fisheries_revenue_", "") %>% 
      str_to_upper()) %>% 
  # filter for not NA
  filter(!is.na(value)) %>% 
  # select columns
  select(time, region, value)
  
# create plot object
p_rgn <- ggplot(
  d_rgn,
  # aesthetics
  aes(
    x     = time, 
    y     = value,
    group = region,
    color = region)) +
  # geometry
  geom_line()

# show plot
p_rgn

# Update labels
p_rgn <- p_rgn +
  labs(
    title  = "Fisheries Revenue",
    x      = "Year",
    y      = "Millions $ (year 2015)",
    color  = "Region")
p_rgn

# Multiple pltos with facet_wrap
p_rgn +
  facet_wrap(vars(region))
```
## Bar plots
```{r}
# Use geom_bar for discrete x data and geom_col for discrete x and continuous y data

library(glue) # for string manipulation
library(lubridate)

yr_max <- year(max(d_rgn$time))

d_rgn %>% 
  # filter by most recent time
  filter(year(time) == yr_max) %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_col() +
  # add labels
  labs(
    title  = glue("Fisheries Revenue for {yr_max}"),
    x      = "Region",
    y      = "Millions $ (year 2015)",
    fill   = "Region")

```
## Boxplots
```{r}
d_rgn %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_boxplot() +
  # add labels
  labs(
    title  = "Fisheries Revenue Variability",
    x      = "Region",
    y      = "Millions $ (year 2015)") +
  # drop legend since redundant with x axis
  theme(
    legend.position = "none")
```

## Violin plots
```{r}
p_rgn_violin <- d_rgn %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_violin() +
  # add labels
  labs(
    title  = "Fisheries Revenue Variability",
    x      = "Region",
    y      = "Millions $ (year 2015)") +
  # drop legend since redundant with x axis
  theme(
    legend.position = "none")
p_rgn_violin
```

# For changing themes

```{r}
p_rgn_violin +
  theme_classic()
```

# For producing interactive plots

## Using `plotly`
```{r}
plotly::ggplotly(p_rgn)
```

## Using `dygraphs` for time series
```{r}
library(dygraphs)

# dygraphs requires data in wide format:
d_rgn_wide <- d_rgn %>% 
  mutate(
    Year = year(time)) %>% 
  select(Year, region, value) %>% 
  pivot_wider(
    names_from  = region,
    values_from = value)
datatable(d_rgn_wide)

d_rgn_wide %>% 
  dygraph() %>% 
  dyRangeSelector()
```

